The Role of Variability in Learning Generalization: A Computational Modeling Approach

### 2024-07-09 Sections/HTW.qmd
update with more of Rob's edits, through end of Project 2 modeling results

*Participants* A total of 156 participants were recruited from the Indiana University Introductory Psychology Course. Participants were randomly assigned to one of two training conditions: varied training or constant training.

*Task.* The "Hit The Wall" (HTW) visuomotor extrapolation task task was programmed in Javascript, making heavy use of the [phaser.io](https://phaser.io/) game library. The HTW task involved launching a projectile such that it would strike the "wall" at target speed indicated at the top of the screen (see @fig-htw-task). The target velocities were given as a range, or band, of acceptable velocity values  (e.g. band 800-1000). During the training stage, participants received feedback indicating whether they had hit the wall within the target velocity band, or how many units their throw was above or below from the target band. Participants were instructed that only the x velocity component of the ball was relevant to the task. The y velocity, or the location at which the ball struck the wall, had no influence on the task feedback.

@@

```{r}
#| label: fig-e1-train-dev
#| fig-cap: Experiment 1 Training Stage. Deviations from target band across training blocks. Lower values represent greater accuracy. 
p1 <- trainE1 |> ggplot(aes(x = Trial_Bin, y = dist, color = condit)) +
    stat_summary(geom = "line", fun = mean) +
    stat_summary(geom = "errorbar", fun.data = mean_se, width = .4, alpha = .7) +
    facet_wrap(~vb)+
    scale_x_continuous(breaks = seq(1, nbins + 1)) +
    theme(legend.title=element_blank()) + 
    labs(y = "Deviation", x="Training Block") 
#ggsave(here("Assets/figs/e1_train_deviation.png"), p1, width = 8, height = 4,bg="white")
p1
```

@@

**Experiment 3 testing discrimination**. Bayesian Mixed Model Predicting Vx as a function of condition (Constant vs. Varied) and Velocity Band. The Intercept represents the baseline condition (constant training & original order), and the Band coefficient represents the  slope for the baseline condition. The interaction terms which include condit and Band (e.g., conditVaried:Band & conditVaried:bandOrderReverse:band) respectively indicate the how the slopes of the varied-original condition differed from the baseline condition, and how varied-reverse condition differed from the varied-original condition {.hover .sm}
:::

@@

```{r}
#| label: fig-e3-test-vx
#| fig-cap: e3 testing x velocities. Translucent bands with dash lines indicate the correct range for each velocity band. 
#| fig-width: 11
#| fig-height: 10
##| column: screen-inset-right
# testE3 |> filter(bandOrder=="Original")|> group_by(id,vb,condit) |> plot_distByCondit()
# testE3 |> filter(bandOrder=="Reverse")|> group_by(id,vb,condit) |> plot_distByCondit() +ggtitle("test")

@@

To estimate the  parameters of ALM and EXAM, we used approximate bayesian computation (ABC), enabling us to obtain an estimate of the posterior distribution of the generalization and learning rate parameters for each individual. ABC belongs to the class of simulation-based inference methods [@cranmerFrontierSimulationbasedInference2020], which have begun being used for parameter estimation in cognitive modeling relatively recently [@turnerTutorialApproximateBayesian2012; @turnerBayesianAnalysisSimulationbased2016; @kangasraasioParameterInferenceComputational2019]. Although they can be applied to any model from which data can be simulated, ABC methods are most useful for complex models that lack an explicit likelihood function (e.g., many neural network models).

The general ABC procedure is to 1) define a prior distribution over model parameters. 2) sample candidate parameter values, $\theta^*$, from the prior. 3) Use $\theta^*$ to generate a simulated dataset, $Data_{sim}$. 4) Compute a measure of discrepancy between the simulated and observed datasets, $discrep$($Data_{sim}$, $Data_{obs}$). 5) Accept $\theta^*$ if the discrepancy is less than the tolerance threshold, $\epsilon$, otherwise reject $\theta^*$. 6) Repeat until desired number of posterior samples are obtained.

@@

The final component of our ABC implementation is the determination of an appropriate value of $\epsilon$. The setting of $\epsilon$ exerts strong influence on the approximated posterior distribution. Smaller values of $\epsilon$ increase the rejection rate, and improve the fidelity of the approximated posterior, while larger values result in an ABC sampler that simply reproduces the prior distribution. Because the individual participants in our dataset differed substantially in terms of the noisiness of their data, we employed an adaptive tolerance setting strategy to tailor $\epsilon$ to each individual. The initial value of $\epsilon$ was set to the overall standard deviation of each individuals velocity values. Thus, sampled parameter values that generated simulated data within a standard deviation of the observed data were accepted, while worse performing parameters were rejected. After every 300 samples the tolerance was allowed to increase only if the current acceptance rate of the algorithm was less than 1%. In such cases, the tolerance was shifted towards the average discrepancy of the 5 best samples obtained thus far. To ensure the acceptance rate did not become overly permissive, $\epsilon$ was also allowed to decrease every time a sample was accepted into the posterior.
:::

@@

The residuals of the model predictions for the testing stage (@fig-htw-resid-pred) also show an unsurprising pattern across fitting methods - with models fit only to the test data showing the best performance, followed by models fit to both training and test data, and with models fit only to the training data showing the worst performance (note that y axes are scaled different between plots). Although EXAM tends to perform better for both Constant and Varied participants (see also @fig-ee-e1), the relative advantage of EXAM is generally larger for the Constant group - a pattern consistent across all three fitting methods. The primary predictive difference between ALM and EXAM is made clear in @fig-cm-vx-pat, which directly compares the observed data against the posterior predictive distributions for both models. Regardless of how the models are fit, only EXAM can capture the pattern where participants are able to discriminate all 6 target bands.

@@

```{r}
#| label: tbl-htw-ee-e23
#| tbl-cap: "Results of Bayesian Regression models predicting model error as a function of Model (ALM vs. EXAM), Condition (Constant vs. Varied), and the interaction between Model and Condition. The values represent the estimate coefficient for each term, with 95% credible intervals in brackets.  The intercept reflects the baseline of ALM and Constant. The other estimates indicate deviations from the baseline for the EXAM mode and varied condition. Lower values indicate better model fit."
#| cache: false

@@

*Model Fits to Experiment 2 and 3.* Data from Experiments 2 and 3 were fit to ALM and EXAM in the same manner as Experiment1 . For brevity, we only plot and discuss the results of the "fit to training and testing data" models - results from the other fitting methods can be found in the appendix. The model fitting results for Experiments 2 and 3 closely mirrored those observed in Experiment 1. The Bayesian regression models predicting model error as a function of Model (ALM vs. EXAM), Condition (Constant vs. Varied), and their interaction (see @tbl-htw-ee-e23) revealed a consistent main effect of Model across all three experiments. The negative coefficients for the ModelEXAM term (Exp 2: $\beta$ = `r bm1_e2$estimate`, 95% CrI `r bm1_e2$conf.low`, `r bm1_e2$conf.high`, pd = `r bm1_e2$pd`; Exp 3: $\beta$ = `r bm1_e3$estimate`, 95% CrI `r bm1_e3$conf.low`, `r bm1_e3$conf.high`, pd = `r bm1_e3$pd`) indicate that EXAM outperformed ALM in both experiments. Furthermore, the interaction between Model and Condition was significant in both Experiment 2 ($\beta$ = `r bm3_e2$estimate`, 95% CrI `r bm3_e2$conf.low`, `r bm3_e2$conf.high`, pd = `r bm3_e2$pd`) and Experiment 3 ($\beta$ = `r bm3_e3$estimate`, 95% CrI `r bm3_e3$conf.low`, `r bm3_e3$conf.high`, pd = `r bm3_e3$pd`), suggesting that the superiority of EXAM over ALM was more pronounced for the Constant group compared to the Varied group, as was the case in Experiment 1. Recall that Experiment 3 included participants in both the original and reverse order conditions - and that this manipulation interacted with the effect of training condition. We thus also controlled for band order in our Bayesian Regression assessing the relative performance of EXAM and ALM in Experiment 3. There was a significant three way interaction between Model, Training Condition, and Band Order ($\beta$ = `r bm4_e3$estimate`, 95% CrI `r bm4_e3$conf.low`, `r bm4_e3$conf.high`, pd = `r bm4_e3$pd`), indicating that the relative advantage of EXAM over ALM was only more pronounced in the original order condition, and not the reverse order condition (see @fig-e2_e3_ae).

```{r}
#| label: fig-e2_e3_ae
#| fig-cap: "Conditional effects of Model (ALM vs EXAM) and Condition (Constant vs. Varied) on Model Error for Experiment 2 and 3 data. Experiment 3 also includes a control for the order of training vs. testing bands (original order vs. reverse order)."
#| fig-height: 6
#| fig-width: 9

@@

The superior performance of EXAM, especially for the constant training groups, may initially seem counterintuitive. One might assume that exposure to multiple, varied examples would be necessary to extract an abstract rule. However, EXAM is not a conventional rule-based model; it does not require the explicit abstraction of a rule. Instead, rule-based responses emerge during the retrieval process. The constant groups' formation of a single, accurate input-output association, combined with the usefulness of the zero point, may have been sufficient for EXAM to capture their performance. A potential concern is that the assumption of participants utilizing the zero point essentially transforms the extrapolation problem into an interpolation problem. However, this concern is mitigated by the consistency of the results across both the original and reversed order conditions (the testing extrapolation bands fall in between the constant training band and the 0 point in experiment 1, but not in experiment 2).

The fits to the individual participants also reveal a number of interesting cases where the models struggle to capture the data (@fig-htw-indv-pred). For example participant 68 exhibits a strong a strong non-monotonicity in the highest velocity band, a pattern which ALM can mimic, but which EXAM cannot capture, given it's to enforce a simple linear relationship between target velocity and response. Participant 70 (lower right corner of @fig-htw-indv-pred) had a roughly parabolic response pattern in their observed data, a pattern which neither model can properly reproduce, but which causes EXAM to perform particularly poorly.

*Modeling Limitations.* The present work compared models based on their ability to predict the observed data, without employing conventional model fit indices such as the Akaike Information Criterion (AIC) or the Bayesian Information Criterion (BIC). These indices, which penalize models based on their number of free parameters, would have been of limited utility in this case, as both ALM and EXAM have two free parameters. However, despite having the same number of free parameters, EXAM could still be considered the more complex model, as it incorporates all the components of ALM plus an additional mechanism for rule-based responding. A more comprehensive model comparison approach might involve performing cross-validation with a held-out subset of the data [@mezzadriHoldoutStrategySelecting2022] or penalizing models based on the range of patterns they can produce [@domeGdistanceComparisonModel2023].

>>>>

*Participants* A total of 156 participants were recruited from Indiana University Introductory Psychology Courses. Participants were randomly assigned to one of two training conditions: varied training or constant training.

*Task.* The "Hit The Wall" (HTW) visuomotor extrapolation task task was programmed in JavaScript, making use of the [phaser.io](https://phaser.io/) game library. The HTW task involved launching a projectile such that it would strike the "wall" at the target speed indicated at the top of the screen (see @fig-htw-task). The target velocities were given as a range, or band, of acceptable velocity values (e.g., band 800-1000). During the training stage, participants received feedback indicating whether they had hit the wall within the target velocity band, or how many units their throw was above or below  the target band. Participants were instructed that only the x velocity component of the ball was relevant to the task. The y velocity, or the location at which the ball struck the wall, had no influence on the task feedback.

@@

```{r}
#| label: fig-e1-train-dev
#| fig-cap: Experiment 1 - Training Stage. Deviations from target band across training blocks. Lower values represent greater accuracy. 
p1 <- trainE1 |> ggplot(aes(x = Trial_Bin, y = dist, color = condit)) +
    stat_summary(geom = "line", fun = mean) +
    stat_summary(geom = "errorbar", fun.data = mean_se, width = .4, alpha = .7) +
    facet_wrap(~vb)+
    scale_x_continuous(breaks = seq(1, nbins + 1)) +
    theme(legend.title=element_blank()) + 
    labs(y = "Deviation", x="Training Block") 
#ggsave(here("Assets/figs/e1_train_deviation.png"), p1, width = 8, height = 4,bg="white")
p1
```

@@

**Experiment 3 testing discrimination**. Bayesian Mixed Model Predicting Vx as a function of condition (Constant vs. Varied) and Velocity Band. The Intercept represents the baseline condition (constant training & original order), and the Band coefficient represents the  slope for the baseline condition. The interaction terms which include condit and Band (e.g., conditVaried:Band & conditVaried:bandOrderReverse:band) respectively indicate how the slopes of the varied-original condition differed from the baseline condition, and how varied-reverse condition differed from the varied-original condition {.hover .sm}
:::

@@

```{r}
#| label: fig-e3-test-vx
#| fig-cap: Experiment 3. Empirical distribution of velocities produced in the testing stage. Translucent bands with dash lines indicate the correct range for each velocity band. 
#| fig-width: 11
#| fig-height: 10
##| column: screen-inset-right
# testE3 |> filter(bandOrder=="Original")|> group_by(id,vb,condit) |> plot_distByCondit()
# testE3 |> filter(bandOrder=="Reverse")|> group_by(id,vb,condit) |> plot_distByCondit() +ggtitle("test")

@@

To estimate the parameters of ALM and EXAM, we used approximate Bayesian computation (ABC), enabling us to obtain an estimate of the posterior distribution of the generalization and learning rate parameters for each individual. ABC belongs to the class of simulation-based inference methods [@cranmerFrontierSimulationbasedInference2020], which have begun being used for parameter estimation in cognitive modeling relatively recently [@turnerTutorialApproximateBayesian2012; @turnerBayesianAnalysisSimulationbased2016; @kangasraasioParameterInferenceComputational2019]. Although they can be applied to any model from which data can be simulated, ABC methods are most useful for complex models that lack an explicit likelihood function (e.g., many neural network models).

The general ABC procedure is to 1) define a prior distribution over model parameters. 2) sample candidate parameter values, $\theta^*$, from the prior. 3) Use $\theta^*$ to generate a simulated dataset, $Data_{sim}$. 4) Compute a measure of discrepancy between the simulated and observed datasets, $discrep$($Data_{sim}$, $Data_{obs}$). 5) Accept $\theta^*$ if the discrepancy is less than the tolerance threshold, $\epsilon$, otherwise reject $\theta^*$. 6) Repeat until the desired number of posterior samples are obtained.

@@

The final component of our ABC implementation is the determination of an appropriate value of $\epsilon$. The setting of $\epsilon$ exerts strong influence on the approximated posterior distribution. Smaller values of $\epsilon$ increase the rejection rate, and improve the fidelity of the approximated posterior, while larger values result in an ABC sampler that simply reproduces the prior distribution. Because the individual participants in our dataset differed substantially in terms of the noisiness of their data, we employed an adaptive tolerance setting strategy to tailor $\epsilon$ to each individual. The initial value of $\epsilon$ was set to the overall standard deviation of each individual's velocity values. Thus, sampled parameter values that generated simulated data within a standard deviation of the observed data were accepted, while worse performing parameters were rejected. After every 300 samples the tolerance was allowed to increase only if the current acceptance rate of the algorithm was less than 1%. In such cases, the tolerance was shifted towards the average discrepancy of the 5 best samples obtained thus far. To ensure the acceptance rate did not become overly permissive, $\epsilon$ was also allowed to decrease every time a sample was accepted into the posterior.
:::

@@

The residuals of the model predictions for the testing stage (@fig-htw-resid-pred) show an unsurprising pattern across fitting methods - with models fit only to the test data showing the best performance, followed by models fit to both training and test data, and with models fit only to the training data showing the worst performance (note that Y-axes are scaled different between plots). Although EXAM tends to perform better for both Constant and Varied participants (see also @fig-ee-e1), the relative advantage of EXAM is generally larger for the Constant group - a pattern consistent across all three fitting methods. The primary predictive difference between ALM and EXAM is made clear in @fig-cm-vx-pat, which directly compares the observed data against the posterior predictive distributions for both models. Regardless of how the models are fit, only EXAM can capture the pattern where participants are able to discriminate all 6 target bands.

@@

```{r}
#| label: tbl-htw-ee-e23
#| tbl-cap: "Results of Bayesian Regression models predicting model error as a function of Model (ALM vs. EXAM), Condition (Constant vs. Varied), and the interaction between Model and Condition. The values represent the estimated coefficient for each term, with 95% credible intervals in brackets.  The intercept reflects the baseline of ALM and Constant. The other estimates indicate deviations from the baseline for the EXAM mode and varied condition. Lower values indicate better model fit."
#| cache: false

@@

*Model Fits to Experiment 2 and 3.*  Data from Experiments 2 and 3 were fit to ALM and EXAM in the same manner as Experiment 1. For brevity, we only plot and discuss the results of the "fit to training and testing data" models - results from the other fitting methods can be found in the appendix. The model fitting results for Experiments 2 and 3 closely mirrored those observed in Experiment 1. The Bayesian regression models predicting model error as a function of Model (ALM vs. EXAM), Condition (Constant vs. Varied), and their interaction (see @tbl-htw-ee-e23) revealed a consistent main effect of Model across all three experiments. The negative coefficients for the ModelEXAM term (Exp 2: $\beta$ = `r bm1_e2$estimate`, 95% CrI `r bm1_e2$conf.low`, `r bm1_e2$conf.high`, pd = `r bm1_e2$pd`; Exp 3: $\beta$ = `r bm1_e3$estimate`, 95% CrI `r bm1_e3$conf.low`, `r bm1_e3$conf.high`, pd = `r bm1_e3$pd`) indicate that EXAM outperformed ALM in both experiments. Furthermore, the interaction between Model and Condition was significant in both Experiment 2 ($\beta$ = `r bm3_e2$estimate`, 95% CrI `r bm3_e2$conf.low`, `r bm3_e2$conf.high`, pd = `r bm3_e2$pd`) and Experiment 3 ($\beta$ = `r bm3_e3$estimate`, 95% CrI `r bm3_e3$conf.low`, `r bm3_e3$conf.high`, pd = `r bm3_e3$pd`), suggesting that the superiority of EXAM over ALM was more pronounced for the Constant group compared to the Varied group, as was the case in Experiment 1. Recall that Experiment 3 included participants in both the original and reverse order conditions - and that this manipulation interacted with the effect of training condition. We thus also controlled for band order in our Bayesian Regression assessing the relative performance of EXAM and ALM in Experiment 3. There was a significant three way interaction between Model, Training Condition, and Band Order ($\beta$ = `r bm4_e3$estimate`, 95% CrI `r bm4_e3$conf.low`, `r bm4_e3$conf.high`, pd = `r bm4_e3$pd`), indicating that the relative advantage of EXAM over ALM was only more pronounced in the original order condition, and not the reverse order condition (see @fig-e2_e3_ae).

```{r}
#| label: fig-e2_e3_ae
#| fig-cap: "Conditional effects of Model (ALM vs EXAM) and Condition (Constant vs. Varied) on Model Error for Experiments 2 and 3 data. Experiment 3 also includes a condition for the order of training vs. testing bands (original order vs. reverse order)."
#| fig-height: 6
#| fig-width: 9

@@

The superior performance of EXAM, especially for the constant training groups, may initially seem counterintuitive. One might assume that exposure to multiple, varied examples would be necessary to extract an abstract rule. However, EXAM is not a conventional rule-based model; it does not require the explicit abstraction of a rule. Instead, rule-based responses emerge during the retrieval process. The constant groups' formation of a single, accurate input-output association, combined with the usefulness of the zero point, seem to have been sufficient for EXAM to capture their performance. A potential concern is that the assumption of participants utilizing the zero point essentially transforms the extrapolation problem into an interpolation problem. However, this concern is mitigated by the consistency of the results across both the original and reversed order conditions (the testing extrapolation bands fall in between the constant training band and the 0 point in experiment 1, but not in experiment 2).

The fits to the individual participants also reveal a number of interesting cases where the models struggle to capture the data (@fig-htw-indv-pred). For example participant 68 exhibits a strong non-monotonicity in the highest velocity band, a pattern which ALM can mimic, but which EXAM cannot capture, given that it enforces a simple linear relationship between target velocity and response. Participant 70 (lower right corner of @fig-htw-indv-pred) had a roughly parabolic response pattern in their observed data, a pattern which neither model can properly reproduce, but which causes EXAM to perform particularly poorly.

*Modeling Limitations.* The present work compared models based on their ability to predict the observed data, without employing conventional model fit indices such as the Akaike Information Criterion (AIC) or the Bayesian Information Criterion (BIC). These indices, which penalize models based on their number of free parameters, would have been of limited utility in the current case, as both ALM and EXAM have two free parameters. However, despite having the same number of free parameters, EXAM could still be considered the more complex model, as it incorporates all the components of ALM plus an additional mechanism for rule-based responding. A more comprehensive model comparison approach might involve performing cross-validation with a held-out subset of the data [@mezzadriHoldoutStrategySelecting2022] or penalizing models based on the range of patterns they can produce [@domeGdistanceComparisonModel2023], under the assumption that more constrained models are more impressive when they do adequately fit a given pattern of results.

### 2024-07-09 Sections/full.qmd
update with more of Rob's edits, through end of Project 2 modeling results

All data processing and statistical analyses were performed in R version 4.32 [@rcoreteamLanguageEnvironmentStatistical2020]. To assess differences between groups, we used Bayesian Mixed Effects Regression. Model fitting was performed with the brms package in R [@burknerBrmsPackageBayesian2017], and descriptive stats and tables were extracted with the BayestestR package [@makowskiBayestestRDescribingEffects2019]. Mixed effects regression enables us to take advantage of partial pooling, simultaneously estimating parameters at the individual and group level. Our use of Bayesian, rather than frequentist methods allows us to directly quantify the uncertainty in our parameter estimates, as well as avoiding convergence issues common to the frequentist analogues of our mixed models.

Each model was set to run with 4 chains, 5000 iterations per chain, with the first 2500 discarded as warmup chains. Rhat values were within an acceptable range, with values \<=1.02 (see appendix for diagnostic plots). We used uninformative priors for the fixed effects of the model (condition and velocity band), and weakly informative Student T distributions for for the random effects. For each model, we report 1) the mean values of the posterior distribution for the parameters of interest, 2) the lower and upper credible intervals (CrI), and the probability of direction value (pd).

@@

: **Statistical Model Specifications**. The specifications for the Bayesian regression models used in the analyses of each of the 3 experiments. Comparisons of accuracy use abosulte deviation as the dependent variable, while comparisons of discrimination use the raw velocities produced by participants as the dependent variable.
:::

@@

In each experiment we compare varied and constant conditions in terms of 1) accuracy in the final training block; 2) testing accuracy as a function of band type (trained vs. extrapolation bands); 3) extent of discrimination between all six testing bands. We quantified accuracy as the absolute deviation between the response velocity and the nearest boundary of the target band. Thus, when the target band was velocity 600-800, throws of 400, 650, and 900 would result in deviation values of 200, 0, and 100, respectively. The degree of discrimination between bands was index by fitting a linear model predicting the response velocity as a function of the target velocity. Participants who reliably discriminated between velocity bands tended to haves slope values \~1, while participants who made throws irrespective of the current target band would have slopes \~0.

@@

: **Experiment 1 - End of training performance**. Comparing final training block accuracy in band common to both groups. The Intercept represents the average of the baseline condition (constant training), and the conditVaried coefficient reflects the difference between the constant and varied groups. A larger positive estimates indicates a greater deviation (lower accuracy) for the varied group. CrI values indicate 95% credible intervals. pd is the probability of direction (the % of the posterior on the same side of 0 as the coefficient estimate). {.sm}
:::

@@

```{r}
#| label: fig-e1-test-dev
#| fig-cap: Experiment 1 Testing Accuracy. A) Empricial Deviations from target band during testing without feedback stage. B) Conditional effect of condition (Constant vs. Varied) and testing band type (trained bands vs. novel extrapolation bands) on testing accuracy. Error bars represent 95% credible intervals.
#| fig-width: 11
#| fig-height: 7

@@

The final component of our ABC implementation is the determination of an appropriate value of $\epsilon$. The setting of $\epsilon$ exerts strong influence on the approximated posterior distribution. Smaller values of $\epsilon$ increase the rejection rate, and improve the fidelity of the approximated posterior, while larger values result in an ABC sampler that simply reproduces the prior distribution. Because the individual participants in our dataset differed substantially in terms of the noisiness of their data, we employed an adaptive tolerance setting strategy to tailor $\epsilon$ to each individual. The initial value of $\epsilon$ was set to the overall standard deviation of each individuals velocity values. Thus, sampled parameter values that generated simulated data within a standard deviation of the observed data were accepted, while worse performing parameters were rejected. After every 300 samples the tolerance was allowed to increase only if the current acceptance rate of the algorithm was less than 1%. In such cases, the tolerance was shifted towards the average discrepancy of the 5 best samples obtained thus far. To ensure the acceptance rate did not become overly permissive, $\epsilon$ was also allowed to decrease every time a sample was accepted into the posterior.
::::

>>>>

All data processing and statistical analyses were performed in R version 4.32 [@rcoreteamLanguageEnvironmentStatistical2020]. To assess differences between groups, we used Bayesian Mixed Effects Regression. Model fitting was performed with the brms package in R [@burknerBrmsPackageBayesian2017], and descriptive stats and tables were extracted with the BayestestR package [@makowskiBayestestRDescribingEffects2019]. Mixed effects regression enables us to take advantage of partial pooling, simultaneously estimating parameters at the individual and group level. Our use of Bayesian, rather than frequentist methods allows us to directly quantify the uncertainty in our parameter estimates, as well as avoid convergence issues common to the frequentist analogues of our mixed models.

Each model was set to run with 4 chains, 5000 iterations per chain, with the first 2500 discarded as warmup chains. Rhat values were within an acceptable range, with values \<=1.02 (see appendix for diagnostic plots). We used uninformative priors for the fixed effects of the model (condition and velocity band), and weakly informative Student T distributions for the random effects. For each model, we report 1) the mean values of the posterior distribution for the parameters of interest, 2) the lower and upper credible intervals (CrI), and the probability of direction value (pd).

@@

: **Statistical Model Specifications**. The specifications for the Bayesian regression models used in the analyses of each of the 3 experiments. Comparisons of accuracy use absolute deviation as the dependent variable, while comparisons of discrimination use the raw velocities produced by participants as the dependent variable.
:::

@@

In each experiment we compare varied and constant conditions in terms of 1) accuracy in the final training block; 2) testing accuracy as a function of  band type (trained vs. extrapolation bands); 3) extent of discrimination between all six testing bands. We quantified accuracy as the absolute deviation between the response velocity and the nearest boundary of the target  band. Thus, when the target band was velocity 600-800, throws of 400, 650, and 900 would result in deviation values of 200, 0, and 100, respectively. The degree of discrimination between bands was measured by fitting a linear model predicting the response velocity as a function of the target velocity. Participants who reliably discriminated between velocity bands tended to have slope values \~1, while participants who made throws irrespective of the current target band would have slopes \~0.

@@

: **Experiment 1 - End of training performance**. Comparing final training block accuracy in the band common to both groups. The Intercept represents the average of the baseline condition (constant training), and the conditVaried coefficient reflects the difference between the constant and varied groups. A larger positive estimates indicates a greater deviation (lower accuracy) for the varied group. CrI values indicate 95% credible intervals. pd is the probability of direction (the % of the posterior on the same side of 0 as the coefficient estimate). {.sm}
:::

@@

```{r}
#| label: fig-e1-test-dev
#| fig-cap: Experiment 1 Testing Accuracy. A) Empiricial Deviations from target band during testing without feedback stage. B) Conditional effect of condition (Constant vs. Varied) and testing band type (trained bands vs. novel extrapolation bands) on testing accuracy. Error bars represent 95% credible intervals.
#| fig-width: 11
#| fig-height: 7

@@

The final component of our ABC implementation is the determination of an appropriate value of $\epsilon$. The setting of $\epsilon$ exerts strong influence on the approximated posterior distribution. Smaller values of $\epsilon$ increase the rejection rate, and improve the fidelity of the approximated posterior, while larger values result in an ABC sampler that simply reproduces the prior distribution. Because the individual participants in our dataset differed substantially in terms of the noisiness of their data, we employed an adaptive tolerance setting strategy to tailor $\epsilon$ to each individual. The initial value of $\epsilon$ was set to the overall standard deviation of each individual's velocity values. Thus, sampled parameter values that generated simulated data within a standard deviation of the observed data were accepted, while worse performing parameters were rejected. After every 300 samples the tolerance was allowed to increase only if the current acceptance rate of the algorithm was less than 1%. In such cases, the tolerance was shifted towards the average discrepancy of the 5 best samples obtained thus far. To ensure the acceptance rate did not become overly permissive, $\epsilon$ was also allowed to decrease every time a sample was accepted into the posterior.
::::
